---
title: "Fitness_Project_V2"
author: "Andrew Pettifor & Yoan Lianto"
format: html
editor: visual
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(RTL)
library(tidyverse)
library(plotly)
library(ggplot2)
library(arrow)
library(patchwork)
library(tidyquant)
library(scales)
library(ggridges)
library(viridis)
library(forcats)
library(scales)
library(lubridate)
library(stats)
library(broom)
library(gt)
library(car)
library(GGally)
library(rlang)
library(glue)
library(purrr)
library(ggfortify)
library(ggpmisc)

```

# Collie vs Hound Fitness Project

```{r include=FALSE}

RTLedu::strava

strava_df <- RTLedu::strava 
  
strava_df
```

```{r include=FALSE}
## Refined Data

##### Here we manually remove any unused variables and data, we could remove these using a package like janitor, but by selecting we can visibily see our removed data using a comparitive name function to the oringal dataset. We now have a cleaner dataset to work with before we begin operating on the data.

#### Additionally We have Changed Distance Meters to KM

#### Chnaged Min and Max Elevation to be MEasured in KM

#### We are removing activities that are exclusive to the "Hound" (Street and Track) (These have been removed to ensure fairness in data)

#### Time is the Same sa Elapsed time
#### Title is useless
```

```{r include=FALSE}
strava_df1 <- strava_df %>% 
  select(Activity.Type,
         Date,
         Distance,
         Calories,
         Avg.HR,
         Max.HR,
         Aerobic.TE,
         Avg.Run.Cadence,
         Max.Run.Cadence,
         Avg.Pace,
         Best.Pace,
         Total.Ascent,
         Total.Descent,
         Avg.Stride.Length,
         Min.Temp,
         Best.Lap.Time,
         Number.of.Laps,
         Max.Temp,
         Moving.Time,
         Elapsed.Time,
         Min.Elevation,
         Max.Elevation,
         Runner) %>% 
  relocate(Runner, .before = Activity.Type) %>% 
  mutate(Distance = ifelse(Distance > 100, Distance / 1000, Distance)) %>% #Converts Distance to KM
  mutate(Max.Elevation = as.numeric(Max.Elevation) / 1000) %>% #Converts Elevation from meters to KM
  mutate(Min.Elevation = as.numeric(Min.Elevation) / 1000) %>%   #Converts Elevation from meters to KM
  separate(Avg.Pace, into = c("Minutes", "Seconds", "Milliseconds"), sep = ":", convert = TRUE) %>%
  mutate(
    Avg.Pace = (Minutes * 60) + Seconds + (Milliseconds / 1000)
  ) %>%
  select(-Minutes, -Seconds, -Milliseconds)
  
strava_df1

strava_df2 <- strava_df1 %>% 
  filter(!Activity.Type %in% c("Track Running", "Street Running")) # Ensures runners will only be compared on the same activities

strava_df2

```

### What is Fitness?

Fitness is a broad term that that can be applied to many fields. To define fitness within the constraints of this assignment, we must specify what fitness is in regard to running. Fitness is the ability to endure long distance running over an extended period of time.

##### Before jumping into who is the fitter runner, there lay some reasoning as to how we choose variables that we chose.

```{r echo=FALSE}
runner_activity <- strava_df2 %>% 
  group_by(Activity.Type, Runner) %>% 
  summarise(Count = n(), .groups = "drop") %>% 
  pivot_wider(names_from = Runner, values_from = Count)

runner_activity_unfiltered <- strava_df %>% 
  group_by(Activity.Type, Runner) %>% 
  summarise(Count = n(), .groups = "drop") %>% 
  pivot_wider(names_from = Runner, values_from = Count)


```

Examining our Data we can identify the following activities per Runner:

```{r echo=FALSE}
runner_activity_unfiltered %>% gt() %>% 
  opt_interactive()
```

We notice that Collie does not participate in "Street Running" and "Track Running" as such we will be excluding these activities from our model to give use the following table:

```{r echo=FALSE}

runner_activity%>% gt() %>% 
  opt_interactive()

```

Now that we know what activities we will be comparing our Runners on, we can now begin to determine what fitness is. Since Trail Running is the most common activity between the two Runners, we will be basing fitness metrics based around the Activity of Trail Running, and will be using their shared 3 activities to measure the "Fitness" of Each.

Measurements of Fitness:

To Account for Elevation of Trail Running: Elevation Adjustment = (Total Ascent + Total Descent) / Distance

### Running Efficiency

Stride Efficiency: (Average Stride Length \* Avg Run Cadence)/ (Average Pace \* (1+ Elevation Adjustment) A more efficient stride will reduce energy expendeture, in theory reducing the calories burned per km. (Cal / Dist)

### Cardiovascular Fitness

VO2 Max: 15.3 \* (Max.HR/ Average HR during sustained run)write Compare runners over time, an increase in VO2 max means improved endurance.

### Overall Speed

Elevation Adjusted Pace = Avg. Pace \* (1+ Elevation Adjustment)

### Improvement over time

Aerobic TE Fitness

```{r echo=FALSE , message=FALSE,warning=FALSE}
#New Data Frame with Measurements

strava_df_r <- strava_df2 %>% 
  mutate(Total.Ascent = as.numeric(Total.Ascent),
         Total.Descent = as.numeric(Total.Descent),
         Avg.HR = as.numeric(Avg.HR),
         Max.HR = as.numeric(Max.HR),
         Avg.Run.Cadence = as.numeric(Avg.Run.Cadence),
         Max.Run.Cadence = as.numeric(Max.Run.Cadence),
         Distance = as.numeric(Distance),
         Calories = as.numeric(Calories),
         )
  
strava_df_r <- strava_df_r %>% 
  mutate(Elevation.Adjustment = ((Total.Ascent + Total.Descent)/1000) / Distance) %>% 
  mutate(VO2_Max = 15.3 * (Max.HR / Avg.HR)) %>% 
  mutate(Stride_Efficiency = (Avg.Stride.Length * Avg.Run.Cadence) / (Avg.Pace * (1 + Elevation.Adjustment))) %>% 
  mutate(Adj_Pace = Avg.Pace * (1 + Elevation.Adjustment))

strava_trails <- strava_df_r %>% 
  filter(Activity.Type == "Trail Running")
```

```{r echo=FALSE , message=FALSE,warning=FALSE}
#My code starts here ********************************************************
votwo <- ggplot(strava_df_r, aes(x = Adj_Pace, y = VO2_Max, color = Runner)) + geom_point() + stat_smooth(method = loess , se = FALSE) +
  labs(title = "Endurance Progress: VO2 max over pace", 
       x = "Pace (m/min)", 
       y = "VO2 max") 

plotly::ggplotly(votwo)
#If VO2 max is higher, it's better
```

Hound appears to have a higher VO2 max across a wider range of paces, particularly at higher speeds while Collie shows a smoother trend but has a declining VO2 max as pace increases.

```{r echo=FALSE, warning=FALSE}
tod <- ggplot(strava_df_r, aes(x = Avg.HR , y = VO2_Max, color = Runner)) +
  geom_point() +
  geom_smooth(method = "loess", se = FALSE) +
  labs(title = "VO2 MAx over Heart rate", 
       x = "Heart rate (bpm)", 
       y = "VO2 Max") 

plotly::ggplotly(tod)
```

Hound has higher average VO2 Max, indicating that his heart is more effective at pumping oxygen to fuel his runs.

```{r echo=FALSE, warning=FALSE}
#Pace over Distance
pod <- ggplot(strava_df_r, aes(x = Distance, y = Adj_Pace, color = Runner)) +
  geom_point() + stat_smooth(method = loess, se = FALSE) +
  scale_color_manual(values = c("Collie" = "blue", "Hound" = "orange")) +
  labs(title = "Pace over Distance", 
       y = "Pace(s/km)", 
       x = "Distance(km)") 
ggplotly(pod)
```

Hound has a higher pace over distance in comparison to Collie. However, this indicates that he is running at a slower rate.

```{r echo=FALSE , message=FALSE,warning=FALSE}
#adjusted pace over heart rate 
flo<- ggplot(strava_df_r, aes(x = Adj_Pace, y = Avg.HR, color = Runner)) +
  geom_point() + stat_smooth(method = loess, se = FALSE) +
  scale_color_manual(values = c("Collie" = "red", "Hound" = "orange")) +
  labs(title = "Heart rate over pace", 
       y = "Average heart rate (bpm)", 
       x = "Average Pace (s/km)") 

plotly::ggplotly(flo)
pod | flo
```

Collie generally has lower heart rate than Hound, despite their effort to run similar distance, meaning Collie does not need to exert as much.

```{r echo=FALSE , message=FALSE,warning=FALSE}
#distance over moving time
wer<- ggplot(strava_df_r, aes(x = Distance, y = Moving.Time, color = Runner)) +
  geom_point() + stat_smooth(method = loess, se = FALSE) +
  scale_color_manual(values = c("Collie" = "blue", "Hound" = "green")) +
  labs(title = "How much time taken to cover distance?", 
       y = "Moving time", 
       x = "Distance") 
plotly::ggplotly(wer)
```

To cover similar distance, Hound takes longer time to reach the distance than Collie.

### **Which runner is the fittest?**

**Hound** appears to have a higher **VO2 max** across a wider range of paces, particularly at higher speeds while **Collie** shows a smoother trend but has a declining VO2 max as pace increases, suggesting potential fatigue or inefficiency at higher speeds.**Hound’s** VO2 max is also more variable, indicating fluctuating performance but a generally higher endurance capacity.

On average, Collie's pace is faster than Hound over the various distance. Looking into their heart rate over their pace of run, Collie has slower heart rate even though he's running at a faster pace than Hound. Ultimately, we look at how much time they took to reach the distance, and Hound is slower to reach compared to Collie.

Overall, **Collie is the fitter runner** because he has outperformed Hound in speed, and endurance.

```{r echo=FALSE, warning=FALSE}
#New Data Frame with Measurements

strava_df_r <- strava_df2 %>% 
  mutate(Total.Ascent = as.numeric(Total.Ascent),
         Total.Descent = as.numeric(Total.Descent),
         Avg.HR = as.numeric(Avg.HR),
         Max.HR = as.numeric(Max.HR),
         Avg.Run.Cadence = as.numeric(Avg.Run.Cadence),
         Max.Run.Cadence = as.numeric(Max.Run.Cadence),
         Distance = as.numeric(Distance),
         Calories = as.numeric(Calories),
         Aerobic.TE= as.numeric(Aerobic.TE)
         )
  
strava_df_r <- strava_df_r %>% 
  mutate(Elevation.Adjustment = ((Total.Ascent + Total.Descent)/1000) / Distance) %>% 
  mutate(VO2_Max = 15.3 * (Max.HR / Avg.HR)) %>% 
  mutate(Stride_Efficiency = (Avg.Stride.Length * Avg.Run.Cadence) / (Avg.Pace * (1 + Elevation.Adjustment))) %>% 
  mutate(Adj_Pace = Avg.Pace * (1 + Elevation.Adjustment))

strava_trails <- strava_df_r %>% 
  filter(Activity.Type == "Trail Running")

Hound_trails <- strava_trails %>% 
  filter(Runner == "Hound")

Collie_trails <- strava_trails %>% 
  filter(Runner == "Collie")

```

### Which individual has improved more in fitness?

```{r echo=FALSE, warning=FALSE}
trails_plot <- strava_trails %>% 
  ggplot(aes(x= Date, y=Adj_Pace, color = Runner )) +
           geom_line(size = 1)+
           facet_wrap(~Runner, ncol = 1, scales = "free")+
           scale_x_date(date_labels = "%Y")+
  scale_color_manual(values = c("Hound" = "orange", "Collie" = "blue")) +
           labs(
             title = "Runner progression over time (Adj.Pace)",
             subtitle = "",
             y = "Adj.Pace (Seconds/KM)",
             x= "Date",
             color = "Runner"
           )

trails_plot %>% ggplotly()

```

#### Univariate

```{r echo=FALSE}

VO2_vs_Runner <- 
  lm(VO2_Max ~ Date * Runner, data = strava_trails)
VO2_vs_Runner %>% glance() %>% 
  select(c(1:2,4:5)) %>% 
  mutate(model = paste0("Is there a significant interaction term that indicates differences in improvement rates.")) %>% 
  select(model, everything()) %>% 
  gt() %>% 
  fmt_percent(columns = !matches("statistical|model"), decimals = 2) %>% 
  fmt_number(columns = matches("statistic"), decimals = 0) %>% 
  fmt_percent(columns = matches("p.value"), decimals = 3)

```

The interaction term (Date \* Runner) is significant, this means one runner is improving faster than the other. (p.value \< 0.05)

### First Vs Last VO2

```{r echo=FALSE}
strava_trails %>%
  group_by(Runner) %>%
  summarize(
    Initial_VO2 = first(VO2_Max),
    Final_VO2 = last(VO2_Max),
    Improvement = (Final_VO2 - Initial_VO2) / Initial_VO2) %>% 
  gt() %>% 
  fmt_percent(columns = matches("Improvement"), decimals = 3)
```

While this appears to be a legitimate form of answer, this is actually overall incorrect. This accounts for only the first and last point in a data set. We want to demonstrate this misatribution error now, so you can see that as we move through our analysis, that using linear regressions to create lines of best fit and analysing slope is a far better option.

### VO2 Max Progression Over Time

```{r echo=FALSE}

vo2_trend_plot <- ggplot(strava_trails, aes(x = Date, y = VO2_Max, color = Runner)) +
  geom_point(alpha = 0.5) +  # Scatter plot of data points
  geom_smooth(method = "lm", se = TRUE, formula = y ~ x, size = 1) +  # Linear regression lines
  labs(title = "VO2 Max Progression Over Time",
       subtitle = "Comparing Fitness Improvement of Collie vs. Hound",
       x = "Date", y = "VO2 Max") +
  scale_color_manual(values = c("Hound" = "orange", "Collie" = "blue"))

vo2_trend_plot %>% ggplotly()

collie_model_vo <- lm(VO2_Max ~ Date, data = Collie_trails)
hound_model_vo <- lm(VO2_Max ~ Date, data = Hound_trails)

collie_slope_vo <- coef(collie_model_vo)[2]
hound_slope_vo <- coef(hound_model_vo)[2]

slope_df_vo <- data.frame(
  Runner = c("Collie", "Hound"),
  Slope = c(collie_slope_vo, hound_slope_vo)
)

slope_df_vo %>% gt()
```

From the following graph and slope results we can see that the Hound is not only a better runner in terms of VO2, but they are also improving their VO2, whereas Collie is actually worsening as time progresses, this can perhaps be attributed to the gap year that Collie has when compared to the Hound.

### VO2 Regression

This regression model helps identify what factors drive VO2 max, one of our key indicators of fitness.

```{r echo=FALSE}

V02_model_fit1 <- 
  lm(VO2_Max ~ Distance, data = strava_trails)
V02_model_fit2 <- 
  lm(VO2_Max ~ Distance + Avg.HR, data = strava_trails)
V02_model_fit3 <-
  lm(VO2_Max ~ Distance + Avg.HR + Elevation.Adjustment, data = strava_trails)
V02_model_fit4 <-
  lm(VO2_Max ~ Distance + Avg.HR + Elevation.Adjustment + Aerobic.TE, data = strava_trails)
V02_model_fit5 <-
  lm(VO2_Max ~ Distance + Avg.HR + Elevation.Adjustment + Aerobic.TE + Stride_Efficiency, data = strava_trails)
V02_model_fit6 <-
  lm(VO2_Max ~ Distance + Avg.HR + Elevation.Adjustment + Aerobic.TE + Stride_Efficiency + Adj_Pace, data = strava_trails)


  V02_model_fit1 %>% glance() %>% 
  add_row(V02_model_fit2 %>% glance()) %>%
  add_row(V02_model_fit3 %>% glance()) %>% 
  add_row(V02_model_fit4 %>% glance()) %>% 
  add_row(V02_model_fit5 %>% glance()) %>%
  add_row(V02_model_fit6 %>% glance()) %>%
  select(c(1:2,4:5)) %>% 
  mutate(model = paste0("Is model ", row_number()," significant to measuring performance?")) %>% 
  select(model, everything()) %>% 
  gt() %>% 
  fmt_percent(columns = !matches("statistical|model"), decimals = 2) %>% 
  fmt_number(columns = matches("statistic"), decimals = 0) %>% 
  fmt_percent(columns = matches("p.value"), decimals = 3)
```

### Testing for Breusch-Pagan (BP) test p-values

```{r, echo=FALSE}
#yoans model diufhidushflisudhfiusdfhviasdufhiadfu
models <- list(V02_model_fit1,V02_model_fit2,V02_model_fit3,V02_model_fit4,V02_model_fit5,V02_model_fit6)

library(car)
library(lmtest)

check_bp <- function(model) {
  return(bptest(model)$p.value)
}

bp <- sapply(models, check_bp)
bp
```

Each model's p-value is greater than 0.05, which indicates that we fail to reject the null hypothesis for all models.This suggests that there is no evidence of heteroscedasticity in any of the models.

### VIF Scores

```{r echo=FALSE}

vif(V02_model_fit5)
    
```

What these outputs tell us:

The final model (Model 5) explains nearly 38% of the variance in VO2 Max using Distance, Avg. HR, Elevation Adjustment, Aerobic TE, and Stride Efficiency.

Aerobic TE and Stride Efficiency are the strongest predictors of VO2 Max

Additionally, the VIF scores of around one indicate a lack of multicolinearity.

Thus We can interperet that if a runner improves their stride and aerobic TE over time, their VO2 max should also improve.

```{r echo=FALSE}

Aerobic.TE_trend_plot <-strava_trails %>%
  ggplot(aes(x = Date, y = Aerobic.TE, color = Runner)) +
  geom_point(alpha = 0.5) +  
  geom_smooth(method = "lm", se = TRUE, formula = y ~ x, size = 1) +  
  labs(title = "Aerobic.TE Progression Over Time",
       subtitle = "Comparing Aerobic.TE Fitness Improvement of Collie vs. Hound",
       x = "Date", y = "VO2 Max") +
  scale_color_manual(values = c("Hound" = "orange", "Collie" = "blue"))

Aerobic.TE_trend_plot %>% ggplotly()


collie_model_ab <- lm(Aerobic.TE ~ Date, data = Collie_trails)
hound_model_ab <- lm(Aerobic.TE ~ Date, data = Hound_trails)

collie_slope_ab <- coef(collie_model_ab)[2]
hound_slope_ab <- coef(hound_model_ab)[2]

slope_df_ab <- data.frame(
  Runner = c("Collie", "Hound"),
  Slope = c(collie_slope_ab, hound_slope_ab)
)

slope_df_ab %>% gt()

```

Here we can see that both the Hound and Collie are improving their Aerobic TE over time. Each continues to work harder, regarding trail running however, Collie is working at an increased rate compared to the Hound. This means that the net benefit of each workout is greater for Collie. Even if only marginally.

```{r echo=FALSE}

strava_trails %>% 
  plot_ly(
    y = ~Aerobic.TE,
    color = ~Runner,
    type = "box",
    colors = c("blue","orange")
  ) %>% 
  layout(
    title = list(text = "Trail Aerobic TE Box Plot"),
    xaxis = list(title = ""),
    yaxis = list(
      title = "Aerobic TE"
      ))

```

The previous point is reinforced by this bar graph, as we can see on average, Collies Aerobic TE has a median of 4.4, while the hound is only 3.5.

```{r echo=FALSE, warning=FALSE}
Hound_trails <- strava_trails %>% 
  filter(Runner == "Hound")

Collie_trails <- strava_trails %>% 
  filter(Runner == "Collie")

p1_hound <-
  ggplot(Hound_trails, aes(x= Date, y = Aerobic.TE, col= Aerobic.TE))+
  geom_point() +
  labs(
    title = "Hound Aerobic.TE"
  )

p2_collie <-
  ggplot(Collie_trails, aes(x= Date, y = Aerobic.TE, col= as.character(Aerobic.TE)))+
  geom_point()+
  labs(
    title = "Collie Aerobic.TE"
  )

p1_hound|p2_collie
```

```{r echo=FALSE, warning=FALSE}

stride_trend_plot <-strava_trails %>%
  ggplot(aes(x = Date, y = Stride_Efficiency, color = Runner)) +
  geom_point(alpha = 0.5) +  
  geom_smooth(method = "lm", se = TRUE, formula = y ~ x, size = 1) +  
  labs(title = "Stride Efficiency Progression Over Time",
       subtitle = "Comparing Stride Efficiency Fitness Improvement of Collie vs. Hound",
       x = "Date", y = "Stride Efficiency") +
  scale_color_manual(values = c("Hound" = "orange", "Collie" = "blue"))

stride_trend_plot %>% ggplotly()

collie_model_eff <- lm(Stride_Efficiency ~ Date, data = Collie_trails)
hound_model_eff <- lm(Stride_Efficiency ~ Date, data = Hound_trails)

collie_slope_eff <- coef(collie_model_eff)[2]
hound_slope_eff <- coef(hound_model_eff)[2]

slope_df_eff <- data.frame(
  Runner = c("Collie", "Hound"),
  Slope = c(collie_slope_eff, hound_slope_eff)
)

slope_df_eff %>% gt()
```

Interpreting the graph we can determine two things. Firstly, Collie has a more efficient stride, allowing for more efficient energy consumption. Second, Collie is activiley improving their stride on trail runs, while the hound has had a slight decline in efficiency over the years.

### Adjusted Pace Progression Over Time (Line)

```{r echo=FALSE}
trails_plot <- strava_trails %>% 
  ggplot(aes(x= Date, y=Adj_Pace, color = Runner )) +
           geom_line(size = 1)+
           facet_wrap(~Runner, ncol = 1, scales = "free")+
           scale_x_date(date_labels = "%Y")+
  scale_color_manual(values = c("Hound" = "orange", "Collie" = "blue")) +
           labs(
             title = "Runner progression over time (Adj.Pace)",
             subtitle = "",
             y = "Adj.Pace (Seconds/KM)",
             x= "Date",
             color = "Runner"
           )

trails_plot %>% ggplotly()


```

### Adjusted Pace Progression Over Time (Linear Regression)

```{r echo=FALSE, warning=FALSE}
adj_pace_trend_plot <- strava_trails %>% 
  ggplot(aes(x = Date, y = Adj_Pace, color = Runner)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = TRUE, formula = y ~ x, size = 1) +
  labs(
    title = "Adjusted Pace Progression Over Time",
    subtitle = "Comparing Pace Improvement of Collie vs. Hound",
    x = "Date", y = "Adjusted Pace (Seconds/KM)"
  ) +
  scale_color_manual(values = c("Hound" = "orange", "Collie" = "blue"))

adj_pace_trend_plot %>% ggplotly()



collie_model_pace <- lm(Adj_Pace ~ Date, data = Collie_trails)
hound_model_pace <- lm(Adj_Pace ~ Date, data = Hound_trails)

collie_slope_pace <- coef(collie_model_pace)[2]
hound_slope_pace <- coef(hound_model_pace)[2]

slope_df_pace <- data.frame(
  Runner = c("Collie", "Hound"),
  Slope = c(collie_slope_pace, hound_slope_pace)
)

slope_df_pace %>% gt()
```

### Final Fitness Improvement Score

```{r echo=FALSE, warning=FALSE}

extract_slope <- function(model) {
  coef(model)[2]  
}

slope_df_prog <- strava_trails %>%
  group_by(Runner) %>%
  summarise(
    VO2_Slope = extract_slope(lm(VO2_Max ~ Date, data = cur_data())),
    Stride_Slope = extract_slope(lm(Stride_Efficiency ~ Date, data = cur_data())),
    AerobicTE_Slope = extract_slope(lm(Aerobic.TE ~ Date, data = cur_data())),
    Pace_Slope = -extract_slope(lm(Adj_Pace ~ Date, data = cur_data())),  # Negative means improvement
  ) %>%
  mutate(
    Overall_Progression_Score = (VO2_Slope + Stride_Slope + AerobicTE_Slope + Pace_Slope) / 4
  )

slope_df_prog %>% gt() %>%
  fmt_number(columns = -Runner, decimals = 5)

ggplot(slope_df_prog, aes(x = Runner, y = Overall_Progression_Score, fill = Runner)) +
  geom_bar(stat = "identity", width = 0.5) +
  geom_text(aes(label = round(Overall_Progression_Score, 3)), 
            nudge_y = ifelse(slope_df_prog$Overall_Progression_Score > 0, 0.01, -0.01),  
            size =5, color = "black") +
  labs(
    title = "Overall Running & Fitness Progression Score (Slope-Based)",
    subtitle = "Higher score means a stronger trend of improvement over time",
    x = "Runner", y = "Improvement Score (Avg. Slope)"
  ) +
  scale_fill_manual(values = c("Hound" = "orange", "Collie" = "blue")) +
  theme_minimal()

```

# Coaching Recommendation

We will choose to coach Collie. In regards to Trail Running, Collie has the better overall fitness and improvement score. While Hound does win in sections like VO2, and does a larger variety of running variants, we are basing our fitness purely around a trail running sample. So to conclude, We choose Collie as our Runner. 
